---
title: "ClinEpi Metrics"
output: html_document
---

```{r setup, include=FALSE, echo=TRUE}
library(knitr)
library(tidyverse)
library(kableExtra)
library(rio)
library(ggplot2)
library(scales)
library(RColorBrewer)
library(forcats)


integer_breaks <- function(n = 5, ...) {
  fxn <- function(x) {
    breaks <- floor(pretty(x, n, ...))
    names(breaks) <- attr(breaks, "labels")
    breaks
  }
  return(fxn)
}

globalUsage <-import("data/globalUsage.csv")
googleScholarCitations <-import("data/googleScholarCitations.csv")
websiteUsage <-import("data/websiteUsage_awstats.csv")
websiteStatsGoogle <- import('data/websiteUsage_googleAnalytics.csv')
studySummary <- import("data/Studies_Summary.csv")
dataAccess <- import("data/dataAccess.csv")

today <- Sys.Date()
todayFormated <- format(today, format="%Y%m%d")
```

##ClinEpiPallet Colors
```{r}

clinEpi_colours = list(
  clinEpiDefault = c("#882255", 
                      "#88ccee",
                      "#999933",
                      "#332288",
                      "#44aa99",
                      "#ddcc77",
                      "#cc6677",
                      "#117733"),
  clinEpiDark = c("#731c48",
                    "#71c2ea",
                    "#85852c",
                    "#2b1c73",
                    "#3c9788",
                    "#d7c462",
                    "#c55266",
                    "#0d6029"))

clinEpiDefaultLst = c("#882255", 
                      "#88ccee",
                      "#999933",
                      "#332288",
                      "#44aa99",
                      "#ddcc77",
                      "#cc6677",
                      "#117733")

ClinEpiColor1st <- "#882255"
ClinEpiColor2nd <-  "#88ccee"

```

Pallet Functions
```{r}
clinEpi_palettes = function(name, n, all_palettes = clinEpi_colours, type = c("discrete", "continuous")) {
  palette = all_palettes[[name]]
  if (missing(n)) {
    n = length(palette)
  }
  type = match.arg(type)
  out = switch(type,
               continuous = grDevices::colorRampPalette(palette)(n),
               discrete = palette[1:n]
  )
  structure(out, name = name, class = "palette")
}

scale_colour_clinEpi_d = function(name) {
  ggplot2::scale_colour_manual(values = clinEpi_palettes(name,
                                                    type = "discrete"))
}

scale_fill_clinEpi_d = function(name) {
  ggplot2::scale_fill_manual(values = clinEpi_palettes(name,
                                                    type = "discrete"))
}
```

##ClinEpiDB Global Usage
#### transform data
```{r globalUsage}
#flip data

globalUsageForPlot <- pivot_longer(globalUsage,!Year, names_to = "region", values_to = "n")
globalUsageForPlot$region <- as.factor(globalUsageForPlot$region)
globalUsageForPlot$region <- factor(globalUsageForPlot$region, levels = c("Americas", "Asia", "Europe", "Africa", "Oceania", "Not set"))


```

#### plot
```{r globalUsagePlot}
globalUsagePlot <- ggplot(globalUsageForPlot, aes(x=Year, y=n, fill=region)) +
    geom_bar(stat = "identity", position = position_stack(reverse = TRUE)) +
    scale_fill_clinEpi_d("clinEpiDefault") +
    scale_y_continuous(expand = c(0,0)) +
    theme_bw() + 
    theme(panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          axis.line = element_line(colour = "black"), 
          text=element_text(color = "black"), 
          plot.title = element_text(hjust = 0.5, size = 16,face="bold"),
          axis.title.x = element_text(size=14),
          axis.text.x = element_text(size=14, face = 'bold'),
          axis.title.y = element_text(size=14),
          axis.text.y = element_text(size=14, face = 'bold'))+
    ylab("Users")+
    guides(fill = guide_legend(reverse=TRUE))+
    ggtitle ("ClinEpiDB global usage (Google Analytics)")

globalUsagePlot  

ggsave(paste0("plots/", todayFormated, "_globalUsagePlot" ,".png", sep = ""), plot = globalUsagePlot)

```

##Plot the number of citiations of ClinEpiDB per year
```{r googleScholarCitations}

citations <- ggplot(googleScholarCitations, aes(x=year, y=citations))+
  geom_point(color = ClinEpiColor1st)+
  geom_line(color = ClinEpiColor1st)+
  theme_bw() +
  theme(panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          axis.line = element_line(colour = "black"), 
          text=element_text(color = "black"), 
          plot.title = element_text(hjust = 0.5, size = 16,face="bold"),
          axis.title.x = element_text(size=14),
          axis.text.x = element_text(size=14, face = 'bold'),
          axis.title.y = element_text(size=14),
          axis.text.y = element_text(size=14, face = 'bold'))+ 
    xlab("Year") +
    ylab("Citations")+
    ggtitle ("ClinEpiDB citations per year (Google Scholar)")

citations

ggsave(paste0("plots/",todayFormated,"_citations",".png", sep = ""), plot = citations)
```

##Plot the cumulative number of citiations of ClinEpiDB
#### transform data
```{r googleScholarCumulative}
#calculate cumulative sum
googleScholarCitations$`Total` <-cumsum(googleScholarCitations[, 'citations'])
googleScholarCitations$`Non-study team`  <-cumsum(googleScholarCitations[,'nonStudyTeam'])

googleScholarCumulativeForPlot <- select(googleScholarCitations, c(year, `Total`, `Non-study team`))

googleScholarCitationsForPlot <- tidyr::pivot_longer(googleScholarCumulativeForPlot,!year, names_to = "type", values_to = "n")


```

#### plot 
```{r googleScholarCumulativePlot}
cumlative <- ggplot(googleScholarCitationsForPlot, aes(x=year, y=n, color = type)) +
  geom_point()+
  geom_line()+
  scale_colour_clinEpi_d("clinEpiDefault") +
  theme_bw() +
  theme(panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          legend.title=element_blank(),
          axis.line = element_line(colour = "black"), 
          text=element_text(color = "black"), 
          plot.title = element_text(hjust = 0.5, size = 16,face="bold"),
          axis.title.x = element_text(size=14),
          axis.text.x = element_text(size=14, face = 'bold'),
          axis.title.y = element_text(size=14),
          axis.text.y = element_text(size=14, face = 'bold')) + 
    guides(fill = guide_legend(byrow = TRUE))+
    xlab("Year") +
    ylab("Citations")+
    ggtitle ("ClinEpiDB cumulative citations (Google Scholar)")

cumlative

ggsave(paste0("plots/",todayFormated,"_cumlativeCitations",".png", sep = ""), plot = cumlative)
```

##Plot the number of unique visitors ClinEpiDB received on average per month each year. Data from AW stats
#### transform data
```{r websiteUsageData}

websiteUsage$UniqueVisitorsPerMonth <- (websiteUsage$UniqueVisitors/websiteUsage$months)

websiteUsage$NumberOfVisitsPerMonth <- (websiteUsage$NumberVisits/websiteUsage$months)

websiteUsage$PagesPerMonth <- (websiteUsage$Pages/websiteUsage$months)

websiteUsage$HitsPerMonth <- (websiteUsage$Hits/websiteUsage$months)

websiteUsage$BandwidthPerMonth <-(websiteUsage$Bandwidth/websiteUsage$months)

maxYear <- max(websiteUsage$Year)

```
#### plot
```{r websiteUsageUsers}

uniqueVisitors <- ggplot(websiteUsage, aes(x=Year, y=UniqueVisitorsPerMonth)) +
  geom_line(color = ClinEpiColor1st)+ 
  geom_point(color = ClinEpiColor1st) + 
  scale_x_continuous(breaks = integer_breaks(), limits = c(NA, NA))+
  scale_y_continuous(limits = c(0,NA))+
  theme_bw() +
  theme(panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          axis.line = element_line(colour = "black"), 
          text=element_text(color = "black"), 
          plot.title = element_text(hjust = 0.5, size = 16,face="bold"),
          axis.title.x = element_text(size=14),
          axis.text.x = element_text(size=14, face = 'bold'),
          axis.title.y = element_text(size=14),
          axis.text.y = element_text(size=14, face = 'bold')) + 
    xlab("Year") +
    ylab("Visitors") +
    ggtitle ("ClinEpiDB unique visitors per month (AW Stats)")
   

uniqueVisitors

ggsave(paste0("plots/",todayFormated,"_uniqueVisitorsAWS",".png", sep = ""), plot = uniqueVisitors)
```

##Plot the number of visits ClinEpiDB received on average per month each year. Data from AW stats
```{r r websiteUsagevisits}

visits <- ggplot(websiteUsage, aes(x=Year, y=NumberOfVisitsPerMonth)) +
 geom_line(color = ClinEpiColor1st)+ 
  geom_point(color = ClinEpiColor1st) + 
  scale_x_continuous(breaks = integer_breaks(), limits = c(NA, NA))+
  scale_y_continuous(limits = c(0, NA))+
  theme_bw() +
  theme(panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          axis.line = element_line(colour = "black"), 
          text=element_text(color = "black"), 
          plot.title = element_text(hjust = 0.5, size = 16,face="bold"),
          axis.title.x = element_text(size=14),
          axis.text.x = element_text(size=14, face = 'bold'),
          axis.title.y = element_text(size=14),
          axis.text.y = element_text(size=14, face = 'bold')) +
    xlab("Year") +
    ylab("Visits") +
    ggtitle ("ClinEpiDB visits per month (AW Stats)")

visits

ggsave(paste0("plots/",todayFormated,"_visitsAWS",".png", sep = ""), plot = visits)
```

##Google Stats - Calculate values
```{r websiteStatsGoogleCalcValues}
websiteStatsGoogle$usersMonth <- (websiteStatsGoogle$users)/(websiteStatsGoogle$months)
websiteStatsGoogle$sessionsMonth <- (websiteStatsGoogle$sessions)/(websiteStatsGoogle$months)
maxYear <- max(websiteStatsGoogle$year)

```

##Google Stats - users/month 
```{r websiteStatsGoogleUsersMonth}

usersGoogle <- ggplot(websiteStatsGoogle, aes(x=year, y=usersMonth)) +
  geom_line(color = ClinEpiColor1st)+ 
  geom_point(color = ClinEpiColor1st) + 
  scale_x_continuous(breaks = integer_breaks())+
  scale_y_continuous(limits = c(0, NA))+
  theme_bw() +
  theme(panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          axis.line = element_line(colour = "black"), 
          text=element_text(color = "black"), 
          plot.title = element_text(hjust = 0.5, size = 16,face="bold"),
          axis.title.x = element_text(size=14),
          axis.text.x = element_text(size=14, face = 'bold'),
          axis.title.y = element_text(size=14),
          axis.text.y = element_text(size=14, face = 'bold')) +
    xlab("Year") +
    ylab("Users") +
    ggtitle ("ClinEpiDB users per month (Google)")

usersGoogle

ggsave(paste0("plots/", todayFormated,"_usersGoogle", ".png", sep = ""), plot = usersGoogle)
```

##Google Stats - sessions/month
```{r websiteStatsGoogleSessionsMonth}

sessionsGoogle <- ggplot(websiteStatsGoogle, aes(x=year, y=sessionsMonth)) +
  geom_line(color = ClinEpiColor1st)+ 
  geom_point(color = ClinEpiColor1st) + 
  scale_x_continuous(breaks = integer_breaks())+
  scale_y_continuous(limits = c(0, NA))+
  theme_bw() +
  theme(panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          axis.line = element_line(colour = "black"), 
          text=element_text(color = "black"), 
          plot.title = element_text(hjust = 0.5, size = 16,face="bold"),
          axis.title.x = element_text(size=14),
          axis.text.x = element_text(size=14, face = 'bold'),
          axis.title.y = element_text(size=14),
          axis.text.y = element_text(size=14, face = 'bold')) +
    xlab("Year") +
    ylab("Sessions") +
    ggtitle ("ClinEpiDB sessions per month (Google)")

sessionsGoogle

ggsave(paste0("plots/",todayFormated,"_sessionsGoogle",".png", sep = ""), plot = sessionsGoogle)
```

##Plot the number of records by disease 
```{r studyMetricsByDisease}

studySummary_byDisease <- select(studySummary, c('Disease', 'Households', 'Participants', "Participant repeated measures", "Samples")) %>% 
    mutate(Disease = recode(Disease, 
                            'Diarrheal disease, Malaria, Respiratory condition' = "Other",
                            'Diarrheal disease, Respiratory condition' = "Other",
                            "N/A" = "Other")) %>%  #recode disease so N/A or multiple disease == other
    pivot_longer(!Disease, 
                 names_to = "typeOfCount", 
                 values_to = "count") %>% 
    mutate(count = recode(count, "N/A" = "")) %>% 
    mutate(count = str_remove_all(count, ",")) %>% 
    mutate(count = as.numeric(count)) %>% #convert count to numeric
    mutate(typeOfCount = factor(typeOfCount , levels = c("Households", "Participants", "Participant repeated measures", "Samples"))) %>% 
    mutate(Disease = factor(Disease, levels = c("Diarrheal disease", "Malaria", "Schistosomiasis", "Other","Respiratory condition", "HIV")))

str_wrap_factor <- function(x) {
  levels(x) <- str_wrap(levels(x),15)
  x
}

```

```{r}
studyMetricsByDiseasePlot <- ggplot(studySummary_byDisease, aes(x=str_wrap_factor(typeOfCount), y=count, fill= Disease)) +
    geom_bar(stat = "identity", position = position_stack(reverse = T)) +
    scale_fill_clinEpi_d("clinEpiDefault") +
    scale_x_discrete(limits=rev)+
    scale_y_continuous(labels = comma, expand = c(0,0))+
    theme_bw() + 
    theme(panel.border = element_blank(), 
                panel.grid.major = element_blank(),
                panel.grid.minor = element_blank(),
                axis.line = element_line(colour = "black"),    
                text=element_text(family="sans"),
                plot.title = element_text(hjust = 0.5, size = 16,face="bold"), 
                legend.title = element_blank(), 
                axis.text.x = element_text(face = 'bold'),
                legend.text = element_text(size = 5), 
                axis.text.y = element_text(face = 'bold'),
                legend.key.size = unit(5, 'mm')) + 
          xlab("") +
          ylab("Records")+
          guides(fill = guide_legend(reverse=TRUE))+
          ggtitle("Record count by disease type")
  

studyMetricsByDiseasePlot <- studyMetricsByDiseasePlot + coord_flip()
studyMetricsByDiseasePlot

ggsave(paste0("plots/",todayFormated,"_studyMetricsByDisease",".png", sep = ""), plot = studyMetricsByDiseasePlot)

```

##Plot the number of records by project
```{r studyMetricsByProject}

studySummary_byProject <-select(studySummary, c('Project name', 'Households', 'Participants', "Participant repeated measures", "Samples")) %>% 
    mutate(`Project name` = recode(`Project name`,"N/A" = "Other")) %>% 
    pivot_longer(!`Project name`, 
                 names_to = "typeOfCount", 
                 values_to = "count") %>% 
    mutate(count = recode(count, "N/A" = "")) %>% 
    mutate(count = str_remove_all(count, ",")) %>% 
    mutate(count = as.numeric(count)) %>%  #convert count to numeric
    mutate(typeOfCount = factor(typeOfCount, levels = c("Households", "Participants", "Participant repeated measures", "Samples")))

```

``` {r}
studyMetricsByProjectPlot <- ggplot(studySummary_byProject, aes(x=str_wrap_factor(typeOfCount), y=count, fill=`Project name`)) +
           geom_bar(stat = "identity", position = position_stack(reverse = T)) +
            scale_fill_clinEpi_d("clinEpiDefault") +
            scale_x_discrete(limits=rev)+
            scale_y_continuous(labels = comma, expand = c(0,0))+
            theme_bw() +
            theme(panel.border = element_blank(), 
                panel.grid.major = element_blank(),
                panel.grid.minor = element_blank(),
                axis.line = element_line(colour = "black"),    
                text=element_text(family="sans"),
                plot.title = element_text(hjust = 0.5, size = 16,face="bold"), 
                legend.title = element_blank(), 
                axis.text.x = element_text(face = 'bold'),
                legend.text = element_text(size = 5), 
                axis.text.y = element_text(face = 'bold'),
                legend.key.size = unit(5, 'mm'))  + 
                xlab("") +
                ylab("Records")+
                guides(fill = guide_legend(reverse=TRUE))+
                ggtitle("Record count by projects")
  

studyMetricsByProjectPlot <- studyMetricsByProjectPlot + coord_flip()
studyMetricsByProjectPlot

ggsave(paste0("plots/",todayFormated,"_studyMetricsByProject",".png", sep = ""), plot = studyMetricsByProjectPlot)

```

##Plot the number of records by study design
```{r studyMetricsByStudyDesign}
studySummary_byStudyDesign <- select(studySummary, c("Study design", 'Households', 'Participants', "Participant repeated measures", "Samples")) %>% 
    mutate(`Study design` = recode(`Study design`, "N/A" = "Other", 
                                    "Non-randomized" = "Other",
                                   "Randomized controlled/clinical trial" = "Randomized controlled trial")) %>% 
    pivot_longer(!`Study design`, 
                 names_to = "typeOfCount", 
                 values_to = "count") %>% 
    mutate(count = recode(count, "N/A" = "")) %>% 
    mutate(count = str_remove_all(count, ",")) %>% 
    mutate(count = as.numeric(count)) %>%  #convert count to numeric
    mutate(typeOfCount = factor(typeOfCount, levels = c("Households", "Participants", "Participant repeated measures", "Samples"))) %>% 
    mutate(`Study design` = factor(`Study design`, levels = c("Cross-sectional", "Cohort", "Case-control", "Randomized controlled trial", "Other")))

```


```{r}
studySummary_byStudyDesignPlot <- ggplot(studySummary_byStudyDesign, aes(x=str_wrap_factor(typeOfCount), y=count, fill=`Study design`)) +
           geom_bar(stat = "identity", position = position_stack(reverse = T)) +
            scale_fill_clinEpi_d("clinEpiDefault") +
            scale_x_discrete(limits=rev)+
            scale_y_continuous(labels = comma, expand = c(0,0))+
            theme_bw() +
            theme(panel.border = element_blank(), 
                panel.grid.major = element_blank(),
                panel.grid.minor = element_blank(),
                axis.line = element_line(colour = "black"),    
                text=element_text(family="sans"),
                plot.title = element_text(hjust = 0.5, size = 16,face="bold"), 
                legend.title = element_blank(), 
                axis.text.x = element_text(face = 'bold'),
                legend.text = element_text(size = 5), 
                axis.text.y = element_text(face = 'bold'),
                legend.key.size = unit(5, 'mm'))  + 
                xlab("") +
                ylab("Records")+
                guides(fill = guide_legend(reverse=TRUE))+
                ggtitle("Record count by study design")
  

studySummary_byStudyDesignPlot <- studySummary_byStudyDesignPlot + coord_flip()
studySummary_byStudyDesignPlot

ggsave(paste0("plots/",todayFormated,"_studyMetricsByStudyDesign",".png", sep = ""), plot = studySummary_byStudyDesignPlot)

```




##Data Access
#### calculate values and transform
```{r dataAccessCumulative}

colnames(dataAccess) <- c("Year", 'Total requests', 'Approved requests', 'Denied requests')

#calculate cumulative sum
dataAccess$`Total cumulative requests` <-cumsum(dataAccess[, 'Total requests'])
dataAccess$`Approved cumulative requests` <-cumsum(dataAccess[, 'Approved requests'])


dataAccessForPlot <- select(dataAccess, c(Year, `Total cumulative requests`, `Approved cumulative requests`)) %>% 
    rename(`Total requests` = `Total cumulative requests`, `Approved requests` = `Approved cumulative requests`)

dataAccessForPlot <- tidyr::pivot_longer(dataAccessForPlot,!Year, names_to = "type", values_to = "n")

```

##Plot data access requests per year
```{r dataAccessPerYear}
dataAccessRequests <- ggplot(dataAccess, aes(x=Year, y=`Total requests`))+
  geom_point(color = ClinEpiColor1st)+
  geom_line(color = ClinEpiColor1st)+
  theme_bw() +
  theme(panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          axis.line = element_line(colour = "black"), 
          text=element_text(color = "black"), 
          plot.title = element_text(hjust = 0.5, size = 16,face="bold"),
          axis.title.x = element_text(size=14),
          axis.text.x = element_text(size=14, face = 'bold'),
          axis.title.y = element_text(size=14),
          axis.text.y = element_text(size=14, face = 'bold'))+ 
    xlab("Year") +
    ylab("Data Requests")+
    ggtitle ("Data access requests per year")

dataAccessRequests

ggsave(paste0("plots/",todayFormated,"_dataAccessRequests",".png", sep = ""), plot = dataAccessRequests)
```

##Plot cumlative requests per year
```{r dataCumAccessPerYear}

cumlativeRequests <- ggplot(dataAccessForPlot, aes(x=Year, y=n, color = type)) +
  geom_point()+
  geom_line()+
  scale_colour_clinEpi_d("clinEpiDefault") +
  theme_bw() +
  theme(panel.border = element_blank(), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(), 
          legend.title=element_blank(),
          axis.line = element_line(colour = "black"), 
          text=element_text(color = "black"), 
          plot.title = element_text(hjust = 0.5, size = 16,face="bold"),
          axis.title.x = element_text(size=14),
          axis.text.x = element_text(size=14, face = 'bold'),
          axis.title.y = element_text(size=14),
          axis.text.y = element_text(size=14, face = 'bold')) + 
    guides(fill = guide_legend(byrow = TRUE), color = guide_legend(reverse=TRUE))+
    xlab("Year") +
    ylab("Requests")+
    ggtitle ("Cumulative data access requests")

cumlativeRequests

ggsave(paste0("plots/",todayFormated,"_cumlativeDataAccessRequests",".png", sep = ""), plot = cumlativeRequests)
```

