---
title: "Dowload file comparision: study's name"
author: "Author's name"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    latex_engine: xelatex
header-includes:
    - \usepackage{caption}
---

```{r setup, include=FALSE}
pacman::p_load(
  "haven",
  "readr",
  "kableExtra",
  "tidyverse" ,
  "labelled",
  "compareDF",
  "arsenal",
  "dplyr",
  "janitor",
  "stringr"
)
```

* Note:
   * d(1,2,3,4,5,6)n represent New download files (as "x" in R result), d(1,2,3,4,5,6)o represent old download files (as "y" in R result).
   * Column's name has been splitted by label (words) and IRI (OntologyID) \

\

There are xx  download files in new folder and xx files in old folder, the details are as follow: \

|Obs   |File's name (New)         |File's name (old) | #obs (new/old)  |#cols (new/old)|
|:---: |:---                      |:---              | :---       |:---   |
|1     |...Communities.txt        | ...Communities.txt | 7960 / 7960 | 5 / 4 | 
|2     |...Households.txt         |...households.txt|  7722 / 21402 | 23 / 34 |
|3     |...ontologyMetadata.txt   |...ontologyMetadata.txt| 94 / 96 | 16 / 15 |  
|4     |...participants.txt       | ...participants.txt| 13754 / 13574 | 14 / 13 |  
|5     |...Participants_rp_measures.txt  |...observations.txt| 40722 / 40722 | 27 / 27 |  
|6     |...Household_repeated_measures.txt    |      | 21402 / - | 12 / - | 
|7     |...RSRC.txt               |...RSRC.txt| 106378 / 40722 | 73 / 74 | 

\
\newpage

## 1. RSRC fiels \

* Brief description: #can replace based on R result...
  * No substantial changes are detected in these two download files.
  * Var "age" and "Country" are changed IRIs, but they have the same values. 
  * New download file contains many NA value. 


```{r, echo=F}
#clear exsited dataset
rm(list=ls()) 
#loading datasets
df1n<-read.delim("/Users/aaronwlsong/Desktop/EDA_ClinEpiWorkflow/2_QA_Checking/7_washb_kenya/new/ISASimple_Gates_WASHb_Kenya_rct_RSRC.txt")
df1o<-read.delim("/Users/aaronwlsong/Desktop/EDA_ClinEpiWorkflow/2_QA_Checking/7_washb_kenya/old/ISASimple_Gates_WASHb_Kenya_rct_RSRC.txt")
```

### Display different col's name in new and old files \

```{r echo=F}
#Compare new and old file' col's name
all_equal(df1n, df1o) # if the two datasets have different rows, it only generates the number of rows for each dataset, otherwise, it shows uncompatible cols. 
compare_df_cols(df1n, df1o) #detect the difference of cols name b/w new and old file
```


### Display the different IRI \
```{r echo=F, include = FALSE, message=F, warning=F}
#retreive IRI from col's name
colnames1<-lst(df1n,df1o) %>%
          map_dfr(~ tibble(name = names(.x), Class = map_chr(.x, class)), 
          .id = 'grp')%>% #extract colnames from new and old files
                  unnest(name) %>%
          separate(name,into=c("label", "IRI"), 
              sep='\\.\\.', convert = TRUE, extra = "merge")

#mannually change unsecessfully retrieved IRIs, this is bc the col's name has spaces, and R converts all spaces and special character "[" as .. or ..., therefore, some IRIs are subset inaccurately.I just referred dataset"colnmaes1" to copy those IRIs and change them manually. Usually it wont have too many. 
colnames1n<-colnames1%>% 
          mutate(
          IRI=case_when(IRI=="days...OBI_0001169."                                                                 ~"OBI_0001169.",
                        IRI=="using.median.weight.and.median.length.or.height..EUPATH_0035071."                    ~"EUPATH_0035071.",
                        IRI=="caregiver.report..EUPATH_0035108."                                                   ~"EUPATH_0035108.",
                        TRUE~IRI)
        )%>%
        mutate(IRI=ifelse(is.na(IRI), label, IRI) ) #fill in col's name for those vars which do not have IRI.

irin1<-colnames1n %>%
      filter(grp=="df1n")%>%
      select(IRI)
irio1<-colnames1n%>%
      filter(grp=="df1o")%>%
      select(IRI)
#compare different IRI
setdiff(irin1,irio1)  #In new file but not in old file  
setdiff(irio1,irin1)  #In old file but not in new file
```


```{r echo=F}
#Step3. Display the different IRI vars and check their value. I just copied the IRI that I want to display from the result of "setdiff(df1, df2).
diffiri1<-colnames1n %>%
  filter(IRI %in% c(
    "Timepoint"	,
    "Household_Observation_Id",
    "Observation_Id",
    "EUPATH_0000579.",
    "ENVO_00000004.",
    ))%>%
  select(grp, label, IRI)%>%
  spread(key=grp,value=IRI)%>%
  print
```

### Check the value for different IRI vars

```{r echo=F}
sumvar<-function(x){
  s<-summary(x)
  return(s)
}
```

**a. var: Country**
```{r echo=F}
sumvar(df1n$Country..ENVO_00000009.)
sumvar(df1o$Country..ENVO_00000004.)
```

**b. var: age** 
```{r echo=F} 
# for continious vars, if the mean, min, max are all same, then I assume they are the same.
sumvar(df1n$Age..days...OBI_0001169.)
sumvar(df1o$Age..days...EUPATH_0000579.)
```
\newpage

