---
title: "Template for Sheena's R functions"
author: "Sheena Tomko"
date: "Aug 31, 2021"
output: html_document
---

# Introduction
The following R markdown file contains information on what each of my functions does, followed by template code you can update and run that calls said functions. To start, run 000a_studyImport_function.R and save the RData file. This will generate a preliminary file that contains the data, a valueMap, and an allVariables file.

## Template 0002a_WL_convert_xlsx_dd_1_sheet

This function will take an xlsx file with multiple sheets and transform it into a single csv file with all rows bound together.

```{r 0002a convert xlsx dd to single sheet}

EXCELDATADICT <- "./data/datadictionary/dataDict.xlsx" #path of the multisheet xlsx data dictionary
OUTPUT <- "./data/datadictionary/dataDict.csv" #where to save the updated single sheet data dictionary

source("/Users/sshahsimpson/Documents/Git/ClinEpiWorkflow/Main/lib/R/0002a_WS_convert_xlsx_dd_1_sheet.R")      #load the function

dataDict <- xl_sheet_dataDict(EXCELDATADICT) #run the function
write.csv(dataDict, OUTPUT, row.names = F) #save updated data dictionary as single csv file
```

## Template 0002_SST_add_data_dictionary

This function:
1) adds data dictionary information to allVariables file, including the codebookDescription, codebookValues, and notes
2) formats the allVariables file to follow the SOP 

Requirements:
1) the data dictionaries must be in a folder on their own
2) they must be in csv, txt or xls format (1 tab only) 
3) they must have the same headers in each file

After running the function, go through the csv file created and manually indicate whether to "keep" or "discard" variables

```{r 0002 add data dictionary}

load("R_output/001.provide_allVars.RData") # Load your allVars R file made with Danica's studyImport script
ALLVARS <- allVars      #use the allVars file in your global environment
FOLDER <- "R_output/dataDictionary"     #folder containing data dictionaries
TYPE <- ".csv"          #must be ".csv",".txt", or ".xls"
VARIABLE <- "Variable name"     #data dictionary column containing variable names
CODEBOOKDESCRIPTION <- "Variable description" #data dictionary column containing variable descriptions
CODEBOOKVALUES <- "Format values" #data dictionary column containing possible values
NOTESFORDL <- "Missing values" #data dictionary column containing anything else you might want to import
OUTPUT <- "./R_output/provide_allVariables_v1.csv" #where to save the updated allVariables file

source("/Users/sshahsimpson/Documents/Git/ClinEpiWorkflow/Main/lib/R/0002_SST_add_data_dictionary.R")      #load the function

allVars2 <- addDD(ALLVARS, FOLDER, TYPE, VARIABLE, CODEBOOKDESCRIPTION, CODEBOOKVALUES, NOTESFORDL) #run the function

# If multiple files contain the same variable, you will get every possible permutation of variable, codebookDescription, codebookValues, etc. The following line of code is a crude way to clean that up.
allVars3 <- distinct(allVars2)
# The following line will remove any rows where there was a line in the data dictionary that did not map to any variables in the allVariables file
allVars4 <- allVars3[is.na(allVars3$dataFile) == F,]

# Save your updated allVariables file. Then go through manually to decide where to "keep" or "discard"
write.csv(allVars4, OUTPUT, row.names = F)
```


## Template 0003_SST_makeVariableMap

This function:
1) removes all variables marked as "discard" but leaves everything 
else
2) adds columns for label, IRI, parentLabel, parentIRI, repeated, is_temporal, mergeKey, unitLabel, unitIRI, is_featured, hidden, scale, defaultDisplayRangeMin, defaultDisplayRangeMax, and defaultBinWidth
3) removes the columns keepDiscard, max, and min to make a starter variableMap file

After running the function, go through the updated file manually and:
1) For known variables - fill out label, IRI, termType, and repeated
2) For new variables - fill out the label, termType, category, parentLabel, parentIRI, and repeated
3) Create a new row for any new multifilters/categories (new parentLabel with no matching parentIRI yet). Place the parentLabel under label and fill out termType, category, parentLabel, and parentIRI to help place the new category in the hierarchy.
4) For all variables, assess whether is_temporal, mergeKey, unitLabel, unitIRI, is_featured, hidden, scale, defaultDisplayRangeMin, defaultDisplayRangeMax, defaultBinWidth, and forceStringType should be completed

for termType: multifilter = multifilter category, value = multifilter value, variable = maps to a variable, category = no data values.

```{r 0003 make variableMap}
source("/Users/sshahsimpson/Documents/Git/ClinEpiWorkflow/Main/lib/R/0003_SST_makeVariableMap.R") #load the function
ALLVARS <- read.csv("./R_output/provide_allVariables_v1.csv", as.is = T) #load the allVariables file where you've marked variables to keep/discard
OUTPUT <- "./R_output/provide_variableMap_v1.csv" #where to save the new variableMap file
makeVariableMap(ALLVARS, OUTPUT) #run the function, which writes a csv file for you
```

## Template 0004_SST_update_variableMap

This function:
1) cleans up and removes NAs from the file
2) removes quotations from around labels and parentLabels that were copy/pasted from Protege
3) Removes all extra columns that don’t belong in the variableMap file that goes to the ontology team

After running the function:
1) decide whether you will be using "variable" or "uniqueVar" to represent variables and update the file manually accordingly
    a) if using uniqueVar, delete the variable column and rename uniqueVar to variable
    b) if using variable, delete the uniqueVar column 
2) send to the ontology team via Git

```{r 0004 update variableMap}
source("/Users/sshahsimpson/Documents/Git/ClinEpiWorkflow/Main/lib/R/0004_SST_update_variableMap.R") #load the function
VARMAP <- read.csv("./R_output/provide_variableMap_v2.csv", as.is = T) #load the variableMap file that you've been editing
OUTPUT <- "./R_output/provide_variableMap.csv" #where to save the updated file
varMap <- updateVariableMap(VARMAP) #run the function

#if removing uniqueVar, use this code
varMap <- varMap %>% 
    select(-uniqueVar)

#if using uniqueVar and removing variable, use this code
varMap <- varMap %>% 
    select(-variable, variable = uniqueVar)

write.csv(varMap, OUTPUT, row.names = F)
```

## Template 0005_SST_update_valueMap

This function:
1) removes discarded variables from the valueMap file 
2) if a variable was marked as boolean by Danica's script, it maps 0 to "No" and 1 to "Yes"
3) removes duplicate rows if values differ only in capitalization

After running the function:
1) update the variable and variable2 columns to both use either "variable" or "uniqueVar" depending on what you used in the variableMap file
2) check in R to see if variables marked as continuous can be dropped
3) Check in R to see if variables marked as >30 categories (potentially free text fields) can be dropped
4) Manually (or in R) update value mapping as needed

```{r 0005 update valueMap}
source("/Users/sshahsimpson/Documents/Git/ClinEpiWorkflow/Main/lib/R/0005_SST_update_valueMap.R") #load function
load(file = "001.provide_allVars.RData") #load the valueMap Danica's script generated
ALLVARS <- read.csv("./R_output/provide_allVariables_v2.csv", as.is = T) #load the allVariables file that indicated which variables to keep/discard
OUTPUT <- "./R_output/provide_valueMap_v1.txt" #indicate where to save the valueMap file
valueMap2 <- updateValueMap(valueMap, ALLVARS) #run function

#if using variable, use this code
valueMap2 <- valueMap2 %>% 
    mutate(variable2 = variable) %>% #create variable2 column
    relocate(variable2, .before = values) %>% #move variable2
    select(-uniqueVar) %>% #remove uniqueVar
    distinct(variable, values, .keep_all = T) #remove duplicate rows if not using uniqueVar

#if using uniqueVar, use this code
valueMap2 <- valueMap2 %>% 
    mutate(variable2 = uniqueVar) %>% #create variable2 column
    relocate(variable2, .before = values) %>% #move variable2
    select(-variable, variable = uniqueVar) #replace variable with uniqueVar

write.table(valueMap2, OUTPUT, sep = "\t", quote = F, row.names = F)
```

